---
title: "Pandemic pressure and product quality. A time series perspective "
author: "Kerlijn Van Assche"
date: "2026-03-17"
output:
  html_document:
    toc: true
    toc_depth: 6
  word_document:
    toc: true
    toc_depth: '6'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

**Aim of the work:** To examine the temporal association between new COVID-19 cases and PQ-MP events and to identify whether increases in COVID-19 cases precede increases in PQ-MP events. The study uses cross-correlation analysis to explore potential lag structures between the two time series.

This document is an appendix to the doctoral thesis of Kerlijn Van Assche.

Full analytical sequence, following each step described in the method section for each time series pair. Please find R script and datasets on GitHub.

This script

-   is only run on datasets that are not too sparse. (\< 50% zero's in the time-points)

-   assesses stationarity of the data before proceeding

-   applies prewhitening before analysing the time-series (takes into account stationarity and is applied to correct for autocorrelation)

    By using the *forecast* library in R, I can calculate the pre-withened *ccf*  (code: [Chapter 8 Correlations and ARIMA \| Time Series Analysis With R](https://nicolarighetti.github.io/Time-Series-Analysis-With-R/correlations-and-arima.html#cross-correlation); explanation: [How Auto ARIMA Works in R](https://mrinalcs.github.io/how-auto-arima-works)).\

## Datasets

India

-   All medical products

-   COVID-19 medicines, Sanitisers and disinfectants

USA

-   All medical products

-   COVID-19 medicines, Sanitisers and disinfectants,COVID-19 diagnostics, Personal Protective Equipment (PPE)

## Sequence of the code for each dataset

-   1/ Data preparation

-   2/ Plotting - visualization of the data

-   3/ Cross-correlations (incl stationarity & autocorrelation)

    -   **Step 0: Define the time series**

    -   **Step1: Stationarity check and differencing\
        **Define if the Y-serie is stationary. If not, define the number of differencing that is required.)

        -   Stationarity testing using the Augmented Dickey-Fuller (ADF) test

        -   Differencing to achieve stationarity

    -   **Step 2: Autocorrelation analysis using ACF and PACF.**\
        Verify the autocorrelation of the stationary data

    -   **Step 3: baseline cross–correlation**

    -   **Step 4. Prewhitened cross-correlation**

        -   A/ **Prewhitening** (filtering and check model fit)\
            Using ARIMA modelling. To isolate genuine cross-correlations

        -   B/ **Cross-correlation analysis using residuals**

# Pre-amble

## Libraries

```{r, results='hide', message=FALSE, echo=FALSE}
library(tidyverse)
library(knitr) 
library(readr)
library(janitor)

library(tseries)
library(forecast) 

```

## Load data

```{r, results='hide', message=FALSE, echo=TRUE}
#COVID-19 data
data_CV19_monthly <- read_csv("Data/data_CV19_monthly_IN-USA.csv")

#Poor-quality medical products data
data_mq <- read_csv("Data/data_mq_IN-USA.csv")
```

## Functions

#### Write a function to check stationarity

```{r, results='hide', message=FALSE, echo=FALSE}
##Function to check stationarity 
check_stationarity <- function(series) {
  result <- adf.test(series)
  
  cat('ADF Statistic:', result$statistic, '\n')
  cat('p-value:', result$p.value, '\n')
  cat('Alternative Hypothesis:', result$alternative, '\n')
  
  if (result$p.value <= 0.05) {
    cat("\033[32mStationary\033[0m\n")
  } else {
    cat("\033[31mNon-stationary\033[0m\n")
  }
}

##Function to check stationarity and perform necessary differencing. 
adf_until_stationary <- function(time_series,
                                 alpha = 0.05,
                                 max_diff = 4) {
  # Capture the name of the object that was passed in
  var_name <- deparse(substitute(time_series))
  # Basic checks
  if (!is.numeric(time_series)) {
    stop("Input must be a numeric vector (time series).")
  }
  time_series <- as.numeric(time_series)
  # ADF on the original series
  adf_initial <- tseries::adf.test(time_series)
  initial_adf_stat <- unname(adf_initial$statistic) 
  initial_p_value  <- adf_initial$p.value
  initial_is_stationary  <- initial_p_value <= alpha
  #Iterative differencing
  diff_order <- 0L
  time_series_diff     <- time_series
  adf_current <- adf_initial

  while (adf_current$p.value > alpha && diff_order < max_diff) {
    diff_order <- diff_order + 1L
    time_series_diff     <- diff(time_series_diff)
    adf_current <- tseries::adf.test(time_series_diff)
  }
  #Final results
  final_adf_stat <- unname(adf_current$statistic)
  final_p_value  <- adf_current$p.value
  final_is_stationary  <- final_p_value <= alpha
  # Return in a tidy-friendly structure 
  tibble(
    var_name = var_name,
    initial_adf_stat = initial_adf_stat,
    initial_p_value  = initial_p_value,
    initial_is_stationary = initial_is_stationary,
    final_adf_stat   = final_adf_stat,
    final_p_value    = final_p_value,
    diff_order       = diff_order,
    final_is_stationary    = final_is_stationary,
    original_series  = list(time_series),
    transformed_series = list(time_series_diff),
    adf_initial_obj  = list(adf_initial)
  )
}

##Function to print result stationarity
print_check_stationarity <- function(stat_check) {
  cat('\n', stat_check$var_name, '\n')
  cat('Results original time-series:\n')
  print(stat_check$adf_initial_obj[[1]])
  cat('Number of differencing required to obtain stationarity:', stat_check$diff_order, '\n')
}
```

#### Load the function for data-preparation subsets

The function returns a dataset that contains information on the PQ-MQ events of interest and on the CV19-variable of interest for a specific geographical region.

```{r, results='hide', message=FALSE, echo=FALSE}
#Source function "process_data_mqcv_countries" to create right subsets of the data
source("Function_process_data_mqcv_countries.R")
```

# A. INDIA: assess cross-correlation

## A1. INDIA: Covid-19 cases & all MQ events

### 1.1. Data preparation

Prepare the dataset for INDIA: All medical products & new CV19 cases.

```{r, echo=FALSE}

joint_data <- process_data_mqcv_countries(
  covid_data = data_CV19_monthly, 
  mq_data = data_mq, 
  inc_countries = "IND",
  exc_countries = NULL,
  product_subset = NULL, #When NULL, all the products are combined
  covid_subset = "new_cases"
  )
```

### 1.2. Plotting

```{r, echo=FALSE}

inc_countries = "India"
product_subset = "All medical products"
covid_subset = "New COVID-19 cases"
 
#Calculate the scale factor dynamically
scale_factor <- max(joint_data$cv19_variable, na.rm = TRUE) / max(joint_data$mq_events, na.rm = TRUE) *0.8 #introducing a "visual reduction factor" so MQ evnts line is smaller relative to CV19

  ### Plot correlations over time
  plot = ggplot(joint_data, aes(x = year_month)) +
    geom_line(aes(y = cv19_variable, color = "cv19_cases", group = 1)) +
    geom_line(aes(y = mq_events * scale_factor, color = "MQ Events", group = 1)) +
    scale_y_continuous(
      name = "cv19_cases",
      sec.axis = sec_axis(~./scale_factor, name = "MQ Events")
    ) +
    labs(title = str_c("CV19 ", str_to_title(gsub("_", " ", covid_subset)), " and MQ Events over time \n", " For ", inc_countries), x = "Date", subtitle = str_c(product_subset)) +
    scale_color_manual("", 
                       breaks = c("cv19_cases", "MQ Events"),
                       values = c("blue", "red")) +
    theme(
      axis.text.x = element_text (angle = 45, hjust = 1, size = 8),
      axis.title.y.right = element_text(color = "red"),
      axis.title.y.left = element_text(color = "blue")
    ) +
    # scale_x_discrete(limits = unique(category_data$year_month)) +
    geom_point(aes(y = mq_events * scale_factor, color = "MQ Events"), 
               size = 2) +
    geom_point(aes(y = cv19_variable, color = "cv19_cases"), 
               size = 2)

  print(plot)
```

### 1.3. Cross-correlations (incl stationarity & autocorrelation)

**Highlights workflow**

-   Need to account for autocorrelation and non-stationarity, which are common sources of spurious correlation.

-   Apply pre-whitening to isolate genuine cross-correlations

-   Validated model assumptions using residual diagnostics and the Ljung-Box test

For the X-series the analysis linked to the time-series itself is only performed once. This X - series (new CV19 cases) is is the same for all the datasets that include India.\

#### Step 0: Define the time series

```{r}
###define the series that needs to be compared
x_series_CV19 <- joint_data$cv19_variable
y_series <- joint_data$mq_events
```

#### Step 1: Stationarity check and differencing

```{r}
##Check stationarity x-series
check_stationarity(x_series_CV19) #non-stationary
#1st time differencing
check_stationarity(diff(x_series_CV19)) #stationary


##Check stationarity y-series
check_stationarity (y_series) #non-stationary
#1st time differencing
check_stationarity(diff(y_series)) #non-stationary
#1st time differencing
check_stationarity(diff(diff(y_series))) #stationary
```

```{r, message=FALSE, echo=TRUE}
##################################
### Information for dataframe  ###
##################################

##Check stationarity x-series
stat_check_x <- adf_until_stationary(x_series_CV19)
print_check_stationarity(stat_check_x)
#Define stationary x_series
x_series_CV19_s <- stat_check_x$transformed_series[[1]]


##Check stationarity y-series
stat_check_y <- adf_until_stationary(y_series)
print_check_stationarity(stat_check_y)
#Define stationary y_series
y_series_s <- stat_check_y$transformed_series[[1]]

```

#### Step 2: Autocorrelation analysis using ACF and PACF

```{r}
##X-series
acf_x <- acf(x_series_CV19_s, main= "New COVID-19 cases")
pacf_x <- pacf(x_series_CV19_s, main= "New COVID-19 cases")

##Y-series
acf_y <- acf(y_series_s, main= product_subset)
pacf_y <- pacf(y_series_s, main= product_subset)

```

#### Step 3: baseline cross-correlation on stationary data

```{r}
###baseline cross-correlation on stationary data
baseline_ccf <- ccf(x_series_CV19_s, y_series_s, na.action = na.omit, 
main = str_c("New COVID-19 cases (x) and ", product_subset, " (y)"), ylab = "baseline CCF")
```

```{r, message=FALSE, echo=FALSE}
##################################
### Information for dataframe  ###
##################################

##Extract the lags that are significant
#significance lines (CCF uses the 95% CI)
sig_line_rccf <- 1.96 / sqrt(baseline_ccf$n.used)

#filter lags above significance level
sig_lags_rccf <- baseline_ccf$acf %>%
  as.numeric() %>%                     
  tibble(acf = .) %>%
  mutate(lag = baseline_ccf$lag %>% as.numeric()) %>% 
  filter(abs(acf) > sig_line_rccf)
sig_lags_rccf

#Range of lags analysed
min_lag <- min(baseline_ccf$lag)
max_lag <- max(baseline_ccf$lag)

#Print results
cat(
  "Significance threshold (95% CI):", sig_line_rccf, "\n",
  "The range of analysis: ", min_lag, " to ", max_lag, "\n",
    "Significant lags:",
  sep = "")

```

#### Step 4: Prewhitened cross-correlation

##### Step 4A: Prewhitening

```{r}
###Steps to apply prewhitning before calculating the cross-correlation

# fit an ARIMA model
x_model <- auto.arima(x_series_CV19)
# keep the residuals ("white noise")
x_residuals_CV19 <- x_model$residuals

#fit the same ARIMA model to the Y-series by using the "Arima" function in forecast
y_model <- Arima(y_series_s, model=x_model)
# keep the residuals
y_filtered <- residuals(y_model)

```

Validate model assumptions using residual diagnostics and the Ljung-Box test

```{r}
##Analyze residuals to evaluate fit of the model - X-series
# acf(x_residuals_CV19)
# pacf(x_residuals_CV19)
# hist(x_residuals_CV19, main = "Histogram of residuals - x series (CV19 cases)")
checkresiduals(x_model)


##Analyze residuals to evaluate fit of the model - Y-series
# acf(y_filtered)
# pacf(y_filtered)
# hist(y_filtered, main = "Histogram of residuals - filtered y series (MQ events")
checkresiduals(y_model)


```

##### Step 4b: Cross-correlation analysis using residuals

Evaluate the correlation between residuals of X- and Y-series using the cross-correlation function

```{r}
# apply the ccf to the residuals
prewhiten_ccf <- ccf(x_residuals_CV19, y_filtered, 
                     main = str_c("Filtered residuals of New COVID-19 cases (x) and ",
                                  product_subset, " (y)"), 
                     ylab = "prewhitened CCF")
```

```{r, message=FALSE, echo=FALSE}
#########################################
### Additional information extraction ###
#########################################

##Extract the lags that are significant
#significance lines (CCF uses the 95% CI)
sig_line_pccf <- 1.96 / sqrt(prewhiten_ccf$n.used)

#filter lags above significance level
sig_lags_pccf <- prewhiten_ccf$acf %>%
  as.numeric() %>%                     
  tibble(acf = .) %>%
  mutate(lag = prewhiten_ccf$lag %>% as.numeric()) %>% 
  filter(abs(acf) > sig_line_pccf)

#Range of lags analysed
min_lag <- min(prewhiten_ccf$lag)
max_lag <- max(prewhiten_ccf$lag)

#Print results
cat(
  "Significance threshold (95% CI):", sig_line_pccf, "\n",
  "The range of analysis: ", min_lag, " to ", max_lag, "\n",
    "Significant lags:",
  sep = ""
)
sig_lags_pccf

```

## A2. INDIA new COVID-19 cases & medicines

### 2.1. Data preparation

Prepare the dataset for INDIA: COVID-19 medicines, new CV19 cases.

```{r, echo=FALSE}
joint_data <- process_data_mqcv_countries(
  covid_data = data_CV19_monthly,
  mq_data = data_mq,
  inc_countries = "IND",
  exc_countries = NULL,
  product_subset = "medicines",
  covid_subset = "new_cases")
```

### 2.2. Plotting

```{r, echo=FALSE}

inc_countries = "India"
product_subset = "COVID-19 medicines"
covid_subset = "New COVID-19 cases"


# ##plot-seperately
# plot(joint_data$year_month, joint_data$cv19_variable, type = "l", col = "blue",xlab = "Date", ylab = "COVID-19 cases", main = "COVID-19 cases Over Time")
#
# plot(joint_data$year_month, joint_data$mq_events, type = "l", col = "red",xlab = "Date", ylab = "MQ Events", main = "MQ Events Over Time")



##Plot together
#Calculate the scale factor dynamically
scale_factor <- max(joint_data$cv19_variable, na.rm = TRUE) / max(joint_data$mq_events, na.rm = TRUE) *0.8 #introducing a "visual reduction factor" so MQ evnts line is smaller relative to CV19

  ### Plot correlations over time
  plot = ggplot(joint_data, aes(x = year_month)) +
    geom_line(aes(y = cv19_variable, color = "cv19_cases", group = 1)) +
    geom_line(aes(y = mq_events * scale_factor, color = "MQ Events", group = 1)) +
    scale_y_continuous(
      name = "cv19_cases",
      sec.axis = sec_axis(~./scale_factor, name = "MQ Events")
    ) +
    labs(title = str_c("CV19 ", str_to_title(gsub("_", " ", covid_subset)), " and MQ Events over time \n", " For ", inc_countries), x = "Date", subtitle = str_c(product_subset)) +
    scale_color_manual("",
                       breaks = c("cv19_cases", "MQ Events"),
                       values = c("blue", "red")) +
    theme(
      axis.text.x = element_text (angle = 45, hjust = 1, size = 8),
      axis.title.y.right = element_text(color = "red"),
      axis.title.y.left = element_text(color = "blue")
    ) +
    # scale_x_discrete(limits = unique(category_data$year_month)) +
    geom_point(aes(y = mq_events * scale_factor, color = "MQ Events"),
               size = 2) +
    geom_point(aes(y = cv19_variable, color = "cv19_cases"),
               size = 2)

  print(plot)
```

### 2.3. Cross-correlations (incl stationarity & autocorrelation)

#### Step 0: Define the time series

```{r}
####define the series that needs to be compared
x_series_CV19 <- joint_data$cv19_variable
y_series <- joint_data$mq_events

```

#### Step 1: Stationarity check and differencing

```{r}
# ##Check stationarity x-series
# check_stationarity(x_series_CV19) #non-stationary
# #1st time differencing
# check_stationarity(diff(x_series_CV19)) #stationary


##Check stationarity y-series
check_stationarity (y_series) #non-stationary
#1st time differencing
check_stationarity(diff(y_series)) #non-stationary
#2nd time differencing
check_stationarity(diff(diff(y_series))) #stationary
```

```{r, message=FALSE, echo=TRUE}
##################################
### Information for dataframe  ###
##################################

# ##Check stationarity x-series
# stat_check_x <- adf_until_stationary(x_series_CV19)
# print_check_stationarity(stat_check_x)
# #Define stationary x_series
# x_series_CV19_s <- stat_check_x$transformed_series[[1]]


##Check stationarity y-series
stat_check_y <- adf_until_stationary(y_series)
print_check_stationarity(stat_check_y)
#Define stationary y_series
y_series_s <- stat_check_y$transformed_series[[1]]
```

#### Step 2: Autocorrelation analysis using ACF and PACF

```{r}
# ##X-series
# acf_x <- acf(x_series_CV19_s, main= "New COVID-19 cases")
# pacf_x <- pacf(x_series_CV19_s, main= "New COVID-19 cases")

##Y-series
acf_y <- acf(y_series_s, main= product_subset)
pacf_y <- pacf(y_series_s, main= product_subset)
```

#### Step 3: baseline cross-correlation on stationary data

```{r}
###baseline cross-correlation on stationary data
baseline_ccf <- ccf(x_series_CV19_s, y_series_s, na.action = na.omit,
main = str_c("New COVID-19 cases (x) and ", product_subset, " (y)"), ylab = "baseline CCF")

baseline_ccf_in_med <- baseline_ccf

```

```{r, message=FALSE, echo=FALSE}
##################################
### Information for dataframe  ###
##################################

##Extract the lags that are significant
#significance lines (CCF uses the 95% CI)
sig_line_rccf <- 1.96 / sqrt(baseline_ccf$n.used)

#filter lags above significance level
sig_lags_rccf <- baseline_ccf$acf %>%
  as.numeric() %>%
  tibble(acf = .) %>%
  mutate(lag = baseline_ccf$lag %>% as.numeric()) %>%
  filter(abs(acf) > sig_line_rccf)
sig_lags_rccf

#Range of lags analysed
min_lag <- min(baseline_ccf$lag)
max_lag <- max(baseline_ccf$lag)

#Print results
cat(
  "Significance threshold (95% CI):", sig_line_rccf, "\n",
  "The range of analysis: ", min_lag, " to ", max_lag, "\n",
    "Significant lags:",
  sep = "")
```

#### Step 4: Prewhitened cross-correlation

##### Step 4A: Prewhitening

```{r}
###Steps to apply prewhitning before calculating the cross-correlation

# # fit an ARIMA model
# x_model <- auto.arima(x_series_CV19)
# # keep the residuals ("white noise")
# x_residuals_CV19 <- x_model$residuals

#fit the same ARIMA model to the Y-series by using the "Arima" function in forecast
y_model <- Arima(y_series_s, model=x_model)
# keep the residuals
y_filtered <- residuals(y_model)

```

```{r}
##Analyze residuals to evaluate fit of the model - X-series
# acf(x_residuals_CV19)
# pacf(x_residuals_CV19)
# hist(x_residuals_CV19, main = "Histogram of residuals - x series (CV19 cases)")
# x_check <- checkresiduals(x_model)
# x_check_in_med <- x_check

##Analyze residuals to evaluate fit of the model - Y-series
# acf(y_filtered)
# pacf(y_filtered)
# hist(y_filtered, main = "Histogram of residuals - filtered y series (MQ events")
checkresiduals(y_model)


```

##### Step 4b: Cross-correlation analysis using residuals

Evaluate the correlation between residuals of X- and Y-series using the cross-correlation function

```{r}
# apply the ccf to the residuals
prewhiten_ccf <- ccf(x_residuals_CV19, y_filtered,
                     main = str_c("Filtered residuals of New COVID-19 cases (x) and ",
                                  product_subset, " (y)"),
                     ylab = "prewhitened CCF")
```

```{r, message=FALSE, echo=FALSE}
#########################################
### Additional information extraction ###
#########################################

##Extract the lags that are significant
#significance lines (CCF uses the 95% CI)
sig_line_pccf <- 1.96 / sqrt(prewhiten_ccf$n.used)

#filter lags above significance level
sig_lags_pccf <- prewhiten_ccf$acf %>%
  as.numeric() %>%
  tibble(acf = .) %>%
  mutate(lag = prewhiten_ccf$lag %>% as.numeric()) %>%
  filter(abs(acf) > sig_line_pccf)

#Range of lags analysed
min_lag <- min(prewhiten_ccf$lag)
max_lag <- max(prewhiten_ccf$lag)

#Print results
cat(
  "Significance threshold (95% CI):", sig_line_pccf, "\n",
  "The range of analysis: ", min_lag, " to ", max_lag, "\n",
    "Significant lags:",
  sep = ""
)
sig_lags_pccf

```

## A3. INDIA Covid-19 cases & sanitisers

### 3.1. Data preparation

Prepare the dataset for INDIA: sanitisers & cv19-cases

```{r, echo=FALSE}
joint_data <- process_data_mqcv_countries(
  covid_data = data_CV19_monthly,
  mq_data = data_mq,
  inc_countries = "IND",
  exc_countries = NULL,
  product_subset = "sanitisers_disinfectants",
  covid_subset = "new_cases"
  )
```

### 3.2. Plotting

```{r, echo=FALSE}


inc_countries = "India"
product_subset = "Sanitisers and disinfectants"
covid_subset = "New COVID-19 cases"


# ##plot-seperately
# plot(joint_data$year_month, joint_data$cv19_variable, type = "l", col = "blue",xlab = "Date", ylab = "COVID-19 cases", main = "COVID-19 cases Over Time")
#
# plot(joint_data$year_month, joint_data$mq_events, type = "l", col = "red",xlab = "Date", ylab = "MQ Events", main = "MQ Events Over Time")



##Plot together
#Calculate the scale factor dynamically
scale_factor <- max(joint_data$cv19_variable, na.rm = TRUE) / max(joint_data$mq_events, na.rm = TRUE) *0.8 #introducing a "visual reduction factor" so MQ evnts line is smaller relative to CV19

  ### Plot correlations over time
  plot = ggplot(joint_data, aes(x = year_month)) +
    geom_line(aes(y = cv19_variable, color = "cv19_cases", group = 1)) +
    geom_line(aes(y = mq_events * scale_factor, color = "MQ Events", group = 1)) +
    scale_y_continuous(
      name = "cv19_cases",
      sec.axis = sec_axis(~./scale_factor, name = "MQ Events")
    ) +
    labs(title = str_c("CV19 ", str_to_title(gsub("_", " ", covid_subset)), " and MQ Events over time \n", " For ", inc_countries), x = "Date", subtitle = str_c(product_subset)) +
    scale_color_manual("",
                       breaks = c("cv19_cases", "MQ Events"),
                       values = c("blue", "red")) +
    theme(
      axis.text.x = element_text (angle = 45, hjust = 1, size = 8),
      axis.title.y.right = element_text(color = "red"),
      axis.title.y.left = element_text(color = "blue")
    ) +
    # scale_x_discrete(limits = unique(category_data$year_month)) +
    geom_point(aes(y = mq_events * scale_factor, color = "MQ Events"),
               size = 2) +
    geom_point(aes(y = cv19_variable, color = "cv19_cases"),
               size = 2)

  print(plot)
```

### 3.3. Cross-correlations (incl stationarity & autocorrelation)

#### Step 0: Define the time series

```{r}
####define the series that needs to be compared
x_series_CV19 <- joint_data$cv19_variable
y_series <- joint_data$mq_events

```

#### Step 1: Stationarity check and differencing

```{r}
# ##Check stationarity x-series
# check_stationarity(x_series_CV19) #non-stationary
# #1st time differencing
# check_stationarity(diff(x_series_CV19)) #stationary


##Check stationarity y-series
check_stationarity (y_series) #non-stationary
#1st time differencing
check_stationarity(diff(y_series)) #stationary
```

```{r, message=FALSE, echo=TRUE}
##################################
### Information for dataframe  ###
##################################

# ##Check stationarity x-series
# stat_check_x <- adf_until_stationary(x_series_CV19)
# print_check_stationarity(stat_check_x)
# #Define stationary x_series
# x_series_CV19_s <- stat_check_x$transformed_series[[1]]


##Check stationarity y-series
stat_check_y <- adf_until_stationary(y_series)
print_check_stationarity(stat_check_y)
#Define stationary y_series
y_series_s <- stat_check_y$transformed_series[[1]]
```

#### Step 2: Autocorrelation analysis using ACF and PACF

```{r}
# ##X-series
# acf_x <- acf(x_series_CV19_s, main= "New COVID-19 cases")
# pacf_x <- pacf(x_series_CV19_s, main= "New COVID-19 cases")

##Y-series
acf_y <- acf(y_series_s, main= product_subset)
pacf_y <- pacf(y_series_s, main= product_subset)
```

#### Step 3: baseline cross-correlation on stationary data

```{r}
###baseline cross-correlation on stationary data
baseline_ccf <- ccf(x_series_CV19_s, y_series_s, na.action = na.omit,
main = str_c("New COVID-19 cases (x) and ", product_subset, " (y)"), ylab = "baseline CCF")

baseline_ccf_in_san <- baseline_ccf
```

```{r, message=FALSE, echo=FALSE}
##################################
### Information for dataframe  ###
##################################

##Extract the lags that are significant
#significance lines (CCF uses the 95% CI)
sig_line_rccf <- 1.96 / sqrt(baseline_ccf$n.used)

#filter lags above significance level
sig_lags_rccf <- baseline_ccf$acf %>%
  as.numeric() %>%
  tibble(acf = .) %>%
  mutate(lag = baseline_ccf$lag %>% as.numeric()) %>%
  filter(abs(acf) > sig_line_rccf)
sig_lags_rccf

#Range of lags analysed
min_lag <- min(baseline_ccf$lag)
max_lag <- max(baseline_ccf$lag)

#Print results
cat(
  "Significance threshold (95% CI):", sig_line_rccf, "\n",
  "The range of analysis: ", min_lag, " to ", max_lag, "\n",
    "Significant lags:",
  sep = "")
```

#### Step 4: Prewhitened cross-correlation

As there is no cross-correlation detected for the baseline CCF, no need to proceed and perform prewhitening.

============================================================

# B. USA

Datasets to look at:

USA

-   all

-   medicines, sanitizers, diagnostics, ppe

## B1. USA: Covid-19 cases & all MQ events

### B1.1. Data preparation

Prepare the dataset for USA , all MQ events combined, cv19-cases

```{r, echo=FALSE}
joint_data <- process_data_mqcv_countries(
  covid_data = data_CV19_monthly,
  mq_data = data_mq,
  inc_countries = "USA",
  exc_countries = NULL,
  product_subset = NULL, #When NULL, all the products are combined
  covid_subset = "new_cases"
  )
```

### B1.2. Plotting

```{r, echo=FALSE}

inc_countries = "United States of America"
product_subset = "All medical products"
covid_subset = "New COVID-19 cases"


# ##plot-seperately
# plot(joint_data$year_month, joint_data$cv19_variable, type = "l", col = "blue",xlab = "Date", ylab = "COVID-19 cases", main = "COVID-19 cases Over Time")
#
# plot(joint_data$year_month, joint_data$mq_events, type = "l", col = "red",xlab = "Date", ylab = "MQ Events", main = "MQ Events Over Time")
#


##Plot together
#Calculate the scale factor dynamically
scale_factor <- max(joint_data$cv19_variable, na.rm = TRUE) / max(joint_data$mq_events, na.rm = TRUE) *0.8 #introducing a "visual reduction factor" so MQ evnts line is smaller relative to CV19

  ### Plot correlations over time
  plot = ggplot(joint_data, aes(x = year_month)) +
    geom_line(aes(y = cv19_variable, color = "cv19_cases", group = 1)) +
    geom_line(aes(y = mq_events * scale_factor, color = "MQ Events", group = 1)) +
    scale_y_continuous(
      name = "cv19_cases",
      sec.axis = sec_axis(~./scale_factor, name = "MQ Events")
    ) +
    labs(title = str_c("CV19 ", str_to_title(gsub("_", " ", covid_subset)), " and MQ Events over time \n", " For ", inc_countries), x = "Date", subtitle = str_c(product_subset)) +
    scale_color_manual("",
                       breaks = c("cv19_cases", "MQ Events"),
                       values = c("blue", "red")) +
    theme(
      axis.text.x = element_text (angle = 45, hjust = 1, size = 8),
      axis.title.y.right = element_text(color = "red"),
      axis.title.y.left = element_text(color = "blue")
    ) +
    # scale_x_discrete(limits = unique(category_data$year_month)) +
    geom_point(aes(y = mq_events * scale_factor, color = "MQ Events"),
               size = 2) +
    geom_point(aes(y = cv19_variable, color = "cv19_cases"),
               size = 2)

  print(plot)
```

### B1.3. Cross-correlations (incl stationarity & autocorrelation)

#### Step 0: Define the time series

```{r}
###define the series that needs to be compared
x_series_CV19 <- joint_data$cv19_variable
y_series <- joint_data$mq_events
```

#### Step 1: Stationarity check and differencing

```{r}
##Check stationarity x-series
check_stationarity(x_series_CV19) #non-stationary
#1st time differencing
check_stationarity(diff(x_series_CV19)) #stationary


##Check stationarity y-series
check_stationarity(y_series) #stationary
```

```{r, message=FALSE, echo=TRUE}
##################################
### Information for dataframe  ###
##################################

##Check stationarity x-series
stat_check_x <- adf_until_stationary(x_series_CV19)
print_check_stationarity(stat_check_x)
#Define stationary x_series
x_series_CV19_s <- stat_check_x$transformed_series[[1]]


##Check stationarity y-series
stat_check_y <- adf_until_stationary(y_series)
print_check_stationarity(stat_check_y)
#Define stationary y_series
y_series_s <- stat_check_y$transformed_series[[1]]
```

#### Step 2: Autocorrelation analysis using ACF and PACF

```{r}
##X-series
acf_x <- acf(x_series_CV19_s, main= "New COVID-19 cases")
pacf_x <- pacf(x_series_CV19_s, main= "New COVID-19 cases")

##Y-series
acf_y <- acf(y_series_s, main= product_subset)
pacf_y <- pacf(y_series_s, main= product_subset)
```

#### Step 3: baseline cross-correlation on stationary data

```{r}
###baseline cross-correlation on stationary data
baseline_ccf <- ccf(x_series_CV19_s, y_series_s, na.action = na.omit,
main = str_c("New COVID-19 cases (x) and ", product_subset, " (y)"), ylab = "baseline CCF")
```

```{r, message=FALSE, echo=FALSE}
##################################
### Information for dataframe  ###
##################################

##Extract the lags that are significant
#significance lines (CCF uses the 95% CI)
sig_line_rccf <- 1.96 / sqrt(baseline_ccf$n.used)

#filter lags above significance level
sig_lags_rccf <- baseline_ccf$acf %>%
  as.numeric() %>%
  tibble(acf = .) %>%
  mutate(lag = baseline_ccf$lag %>% as.numeric()) %>%
  filter(abs(acf) > sig_line_rccf)
sig_lags_rccf


#Range of lags analysed
min_lag <- min(baseline_ccf$lag)
max_lag <- max(baseline_ccf$lag)

#Print results
cat(
  "Significance threshold (95% CI):", sig_line_rccf, "\n",
  "The range of analysis: ", min_lag, " to ", max_lag, "\n",
    "Significant lags:",
  sep = "")
```

#### Step 4: Prewhitened cross-correlation

As there is no cross-correlation detected for the baseline CCF, no need to proceed and perform prewhitening.

## B2. USA: Covid-19 cases & medicines

### B2.1. Data preparation

Prepare the dataset for USA, medicines, CV19 cases

```{r, echo=FALSE}
joint_data <- process_data_mqcv_countries(
  covid_data = data_CV19_monthly,
  mq_data = data_mq,
  inc_countries = "USA",
  exc_countries = NULL,
  product_subset = "medicines",
  covid_subset = "new_cases")
```

### B2.2. Plotting

```{r,echo=FALSE}
inc_countries = "United States of America"
product_subset = "COVID-19 medicines"
covid_subset = "New COVID-19 cases"


# ##plot-seperately
# plot(joint_data$year_month, joint_data$cv19_variable, type = "l", col = "blue",xlab = "Date", ylab = "COVID-19 cases", main = "COVID-19 cases Over Time")
#
# plot(joint_data$year_month, joint_data$mq_events, type = "l", col = "red",xlab = "Date", ylab = "MQ Events", main = "MQ Events Over Time")



##Plot together
#Calculate the scale factor dynamically
scale_factor <- max(joint_data$cv19_variable, na.rm = TRUE) / max(joint_data$mq_events, na.rm = TRUE) *0.8 #introducing a "visual reduction factor" so MQ evnts line is smaller relative to CV19

  ### Plot correlations over time
  plot = ggplot(joint_data, aes(x = year_month)) +
    geom_line(aes(y = cv19_variable, color = "cv19_cases", group = 1)) +
    geom_line(aes(y = mq_events * scale_factor, color = "MQ Events", group = 1)) +
    scale_y_continuous(
      name = "cv19_cases",
      sec.axis = sec_axis(~./scale_factor, name = "MQ Events")
    ) +
    labs(title = str_c("CV19 ", str_to_title(gsub("_", " ", covid_subset)), " and MQ Events over time \n", " For ", inc_countries), x = "Date", subtitle = str_c(product_subset)) +
    scale_color_manual("",
                       breaks = c("cv19_cases", "MQ Events"),
                       values = c("blue", "red")) +
    theme(
      axis.text.x = element_text (angle = 45, hjust = 1, size = 8),
      axis.title.y.right = element_text(color = "red"),
      axis.title.y.left = element_text(color = "blue")
    ) +
    # scale_x_discrete(limits = unique(category_data$year_month)) +
    geom_point(aes(y = mq_events * scale_factor, color = "MQ Events"),
               size = 2) +
    geom_point(aes(y = cv19_variable, color = "cv19_cases"),
               size = 2)

  print(plot)
```

### B2.3. Cross-correlations (incl stationarity & autocorrelation)

#### Step 0: Define the time series

```{r}
####define the series that needs to be compared
x_series_CV19 <- joint_data$cv19_variable
y_series <- joint_data$mq_events

```

#### Step 1: Stationarity check and differencing

```{r}
##Check stationarity x-series
check_stationarity(x_series_CV19) #non-stationary
#1st time differencing
check_stationarity(diff(x_series_CV19)) #stationary

##Check stationarity y-series
check_stationarity (y_series) #non-stationary
#1st time differencing
check_stationarity(diff(y_series)) #non-stationary
#2nd time differencing
check_stationarity(diff(diff(y_series))) #stationary

```

```{r, message=FALSE, echo=TRUE}
##################################
### Information for dataframe  ###
##################################

# ##Check stationarity x-series
# stat_check_x <- adf_until_stationary(x_series_CV19)
# print_check_stationarity(stat_check_x)
# #Define stationary x_series
# x_series_CV19_s <- stat_check_x$transformed_series[[1]]


##Check stationarity y-series
stat_check_y <- adf_until_stationary(y_series)
print_check_stationarity(stat_check_y)
#Define stationary y_series
y_series_s <- stat_check_y$transformed_series[[1]]
```

#### Step 2: Autocorrelation analysis using ACF and PACF

```{r}
# ##X-series
# acf_x <- acf(x_series_CV19_s, main= "New COVID-19 cases")
# pacf_x <- pacf(x_series_CV19_s, main= "New COVID-19 cases")

##Y-series
acf_y <- acf(y_series_s, main= product_subset)
pacf_y <- pacf(y_series_s, main= product_subset)
```

#### Step 3: baseline cross-correlation on stationary data

```{r}
###baseline cross-correlation on stationary data
baseline_ccf <- ccf(x_series_CV19_s, y_series_s, na.action = na.omit,
main = str_c("New COVID-19 cases (x) and ", product_subset, " (y)"), ylab = "baseline CCF")
```

```{r, message=FALSE, echo=FALSE}
##################################
### Information for dataframe  ###
##################################

##Extract the lags that are significant
#significance lines (CCF uses the 95% CI)
sig_line_rccf <- 1.96 / sqrt(baseline_ccf$n.used)

#filter lags above significance level
sig_lags_rccf <- baseline_ccf$acf %>%
  as.numeric() %>%
  tibble(acf = .) %>%
  mutate(lag = baseline_ccf$lag %>% as.numeric()) %>%
  filter(abs(acf) > sig_line_rccf)
sig_lags_rccf

#Range of lags analysed
min_lag <- min(baseline_ccf$lag)
max_lag <- max(baseline_ccf$lag)

#Print results
cat(
  "Significance threshold (95% CI):", sig_line_rccf, "\n",
  "The range of analysis: ", min_lag, " to ", max_lag, "\n",
    "Significant lags:",
  sep = "")

```

#### Step 4: Prewhitened cross-correlation

As there is no cross-correlation detected for the baseline CCF, no need to proceed and perform prewhitening.

## B3. USA: Covid-19 cases & sanitizers

### B3.1. Data preparation

Prepare the dataset for USA, hand-sanitizers, CV19 cases

```{r, echo=FALSE}
joint_data <- process_data_mqcv_countries(
  covid_data = data_CV19_monthly,
  mq_data = data_mq,
  inc_countries = "USA",
  exc_countries = NULL,
  product_subset = "sanitisers_disinfectants",
  covid_subset = "new_cases")
```

### B3.2. Plotting

```{r, echo=FALSE}
inc_countries = "United States of America"
product_subset = "Sanitisers and disinfectants"
covid_subset = "New COVID-19 cases"


##plot-seperately
# plot(joint_data$year_month, joint_data$cv19_variable, type = "l", col = "blue",xlab = "Date", ylab = "COVID-19 cases", main = "COVID-19 cases Over Time")
#
# plot(joint_data$year_month, joint_data$mq_events, type = "l", col = "red",xlab = "Date", ylab = "MQ Events", main = "MQ Events Over Time")



##Plot together
#Calculate the scale factor dynamically
scale_factor <- max(joint_data$cv19_variable, na.rm = TRUE) / max(joint_data$mq_events, na.rm = TRUE) *0.8 #introducing a "visual reduction factor" so MQ evnts line is smaller relative to CV19

  ### Plot correlations over time
  plot = ggplot(joint_data, aes(x = year_month)) +
    geom_line(aes(y = cv19_variable, color = "cv19_cases", group = 1)) +
    geom_line(aes(y = mq_events * scale_factor, color = "MQ Events", group = 1)) +
    scale_y_continuous(
      name = "cv19_cases",
      sec.axis = sec_axis(~./scale_factor, name = "MQ Events")
    ) +
    labs(title = str_c("CV19 ", str_to_title(gsub("_", " ", covid_subset)), " and MQ Events over time \n", " For ", inc_countries), x = "Date", subtitle = str_c(product_subset)) +
    scale_color_manual("",
                       breaks = c("cv19_cases", "MQ Events"),
                       values = c("blue", "red")) +
    theme(
      axis.text.x = element_text (angle = 45, hjust = 1, size = 8),
      axis.title.y.right = element_text(color = "red"),
      axis.title.y.left = element_text(color = "blue")
    ) +
    # scale_x_discrete(limits = unique(category_data$year_month)) +
    geom_point(aes(y = mq_events * scale_factor, color = "MQ Events"),
               size = 2) +
    geom_point(aes(y = cv19_variable, color = "cv19_cases"),
               size = 2)

  print(plot)
```

### B3.3. Cross-correlations (incl stationarity & autocorrelation)

#### Step 0: Define the time series

```{r}
####define the series that needs to be compared
x_series_CV19 <- joint_data$cv19_variable
y_series <- joint_data$mq_events

```

#### Step 1: Stationarity check and differencing

```{r}
# ##Check stationarity x-series
# check_stationarity(x_series_CV19) #non-stationary
# #1st time differencing
# check_stationarity(diff(x_series_CV19)) #stationary

##Check stationarity y-series
check_stationarity (y_series) #non-stationary
#1st time differencing
check_stationarity(diff(y_series)) #non-stationary
#2nd time differencing
check_stationarity(diff(diff(y_series))) #stationary

```

```{r, results='hide', message=FALSE, echo=FALSE}
##################################
### Information for dataframe  ###
##################################

# ##Check stationarity x-series
# stat_check_x <- adf_until_stationary(x_series_CV19)
# print_check_stationarity(stat_check_x)
# #Define stationary x_series
# x_series_CV19_s <- stat_check_x$transformed_series[[1]]


##Check stationarity y-series
stat_check_y <- adf_until_stationary(y_series)
print_check_stationarity(stat_check_y)
#Define stationary y_series
y_series_s <- stat_check_y$transformed_series[[1]]
```

#### Step 2: Autocorrelation analysis using ACF and PACF

```{r}
# ##X-series
# acf_x <- acf(x_series_CV19_s, main= "New COVID-19 cases")
# pacf_x <- pacf(x_series_CV19_s, main= "New COVID-19 cases")

##Y-series
acf_y <- acf(y_series_s, main= product_subset)
pacf_y <- pacf(y_series_s, main= product_subset)
```

#### Step 3: baseline cross-correlation on stationary data

```{r}
###baseline cross-correlation on stationary data
baseline_ccf <- ccf(x_series_CV19_s, y_series_s, na.action = na.omit,
main = str_c("New COVID-19 cases (x) and ", product_subset, " (y)"), ylab = "baseline CCF")
```

```{r, message=FALSE, echo=FALSE}
##################################
### Information for dataframe  ###
##################################

##Extract the lags that are significant
#significance lines (CCF uses the 95% CI)
sig_line_rccf <- 1.96 / sqrt(baseline_ccf$n.used)

#filter lags above significance level
sig_lags_rccf <- baseline_ccf$acf %>%
  as.numeric() %>%
  tibble(acf = .) %>%
  mutate(lag = baseline_ccf$lag %>% as.numeric()) %>%
  filter(abs(acf) > sig_line_rccf)
sig_lags_rccf

#Range of lags analysed
min_lag <- min(baseline_ccf$lag)
max_lag <- max(baseline_ccf$lag)

#Print results
cat(
  "Significance threshold:", sig_line_rccf, "\n",
  "The range of analysis: ", min_lag, " to ", max_lag, "\n",
    "Significant lags:",
  sep = "")

```

```{r}
#########################################
### Additional information extraction ###
#########################################

##Extract the lags that are significant
#significance lines (CCF uses the 95% CI)
sig_line_pccf <- 1.96 / sqrt(prewhiten_ccf$n.used)

#filter lags above significance level
sig_lags_pccf <- prewhiten_ccf$acf %>%
  as.numeric() %>%
  tibble(acf = .) %>%
  mutate(lag = prewhiten_ccf$lag %>% as.numeric()) %>%
  filter(abs(acf) > sig_line_pccf)

#Range of lags analysed
min_lag <- min(prewhiten_ccf$lag)
max_lag <- max(prewhiten_ccf$lag)

#Print results
cat(
  "Significance threshold (95% CI):", sig_line_pccf, "\n",
  "The range of analysis: ", min_lag, " to ", max_lag, "\n",
    "Significant lags:",
  sep = ""
)
sig_lags_pccf

```

#### Step 4: Prewhitened cross-correlation

As there is no cross-correlation detected for the baseline CCF, no need to proceed and perform prewhitening.

```{r, results='hide', message=FALSE, echo=FALSE}
sig_lags_pccf <- "Not applicable"
```

**Conclusion** There is no cross-correlation between the differenced x-series (cv19 cases) and the differenced y-series (MQ events for sanitizers) in the USA.

## B4. USA: Covid-19 cases & diagnostics

### B4.1. Data preparation

Prepare the dataset for USA, diagnostics , CV19 cases

```{r, echo=FALSE}
joint_data <- process_data_mqcv_countries(
  covid_data = data_CV19_monthly,
  mq_data = data_mq,
  inc_countries = "USA",
  exc_countries = NULL,
  product_subset = "diagnostics", #When NULL, all the products are combined
  covid_subset = "new_cases"
  )
```

### B4.2. Plotting

```{r, echo=FALSE}

inc_countries = "United States of America"
product_subset = "COVID-19 diagnostics"
covid_subset = "New COVID-19 cases"


# ##plot-seperately
# plot(joint_data$year_month, joint_data$cv19_variable, type = "l", col = "blue",xlab = "Date", ylab = "COVID-19 cases", main = "COVID-19 cases Over Time")
#
# plot(joint_data$year_month, joint_data$mq_events, type = "l", col = "red",xlab = "Date", ylab = "MQ Events", main = "MQ Events Over Time")



##Plot together
#Calculate the scale factor dynamically
scale_factor <- max(joint_data$cv19_variable, na.rm = TRUE) / max(joint_data$mq_events, na.rm = TRUE) *0.8 #introducing a "visual reduction factor" so MQ evnts line is smaller relative to CV19

  ### Plot correlations over time
  plot = ggplot(joint_data, aes(x = year_month)) +
    geom_line(aes(y = cv19_variable, color = "cv19_cases", group = 1)) +
    geom_line(aes(y = mq_events * scale_factor, color = "MQ Events", group = 1)) +
    scale_y_continuous(
      name = "cv19_cases",
      sec.axis = sec_axis(~./scale_factor, name = "MQ Events")
    ) +
    labs(title = str_c("CV19 ", str_to_title(gsub("_", " ", covid_subset)), " and MQ Events over time \n", " For ", inc_countries), x = "Date", subtitle = str_c(product_subset)) +
    scale_color_manual("",
                       breaks = c("cv19_cases", "MQ Events"),
                       values = c("blue", "red")) +
    theme(
      axis.text.x = element_text (angle = 45, hjust = 1, size = 8),
      axis.title.y.right = element_text(color = "red"),
      axis.title.y.left = element_text(color = "blue")
    ) +
    # scale_x_discrete(limits = unique(category_data$year_month)) +
    geom_point(aes(y = mq_events * scale_factor, color = "MQ Events"),
               size = 2) +
    geom_point(aes(y = cv19_variable, color = "cv19_cases"),
               size = 2)

  print(plot)
```

### B4.3. Cross-correlations (incl stationarity & autocorrelation)

#### Step 0: Define the time series

```{r}
####define the series that needs to be compared
x_series_CV19 <- joint_data$cv19_variable
y_series <- joint_data$mq_events

```

#### Step 1: Stationarity check and differencing

```{r}
##Check stationarity x-series
check_stationarity(x_series_CV19) #non-stationary
#1st time differencing
check_stationarity(diff(x_series_CV19)) #stationary

##Check stationarity y-series
check_stationarity (y_series) #non-stationary
#1st time differencing
check_stationarity(diff(y_series)) #non-stationary
#2nd time differencing
check_stationarity(diff(diff(y_series))) #stationary

```

```{r, message=FALSE, echo=TRUE}
##################################
### Information for dataframe  ###
##################################

# ##Check stationarity x-series
# stat_check_x <- adf_until_stationary(x_series_CV19)
# print_check_stationarity(stat_check_x)
# #Define stationary x_series
# x_series_CV19_s <- stat_check_x$transformed_series[[1]]


##Check stationarity y-series
stat_check_y <- adf_until_stationary(y_series)
print_check_stationarity(stat_check_y)
#Define stationary y_series
y_series_s <- stat_check_y$transformed_series[[1]]
```

#### Step 2: Autocorrelation analysis using ACF and PACF

```{r}
# ##X-series
# acf_x <- acf(x_series_CV19_s, main= "New COVID-19 cases")
# pacf_x <- pacf(x_series_CV19_s, main= "New COVID-19 cases")

##Y-series
acf_y <- acf(y_series_s, main= product_subset)
pacf_y <- pacf(y_series_s, main= product_subset)
```

#### Step 3: baseline cross-correlation on stationary data

```{r}
###baseline cross-correlation on stationary data
baseline_ccf <- ccf(x_series_CV19_s, y_series_s, na.action = na.omit,
main = str_c("New COVID-19 cases (x) and ", product_subset, " (y)"), ylab = "baseline CCF")
```

```{r, message=FALSE, echo=FALSE}
##################################
### Information for dataframe  ###
##################################

##Extract the lags that are significant
#significance lines (CCF uses the 95% CI)
sig_line_rccf <- 1.96 / sqrt(baseline_ccf$n.used)

#filter lags above significance level
sig_lags_rccf <- baseline_ccf$acf %>%
  as.numeric() %>%
  tibble(acf = .) %>%
  mutate(lag = baseline_ccf$lag %>% as.numeric()) %>%
  filter(abs(acf) > sig_line_rccf)
sig_lags_rccf


#Range of lags analysed
min_lag <- min(baseline_ccf$lag)
max_lag <- max(baseline_ccf$lag)

#Print results
cat(
  "Significance threshold (95% CI):", sig_line_rccf, "\n",
  "The range of analysis: ", min_lag, " to ", max_lag, "\n",
    "Significant lags:",
  sep = "")
```

#### Step 4: Prewhitened cross-correlation

As there is no cross-correlation detected for the baseline CCF, no need to proceed and perform prewhitening.

## B5. USA: Covid-19 cases & ppe

Prepare the dataset for USA, PPE (personal protective equipment), CV19 cases

### B5.1. Data preparation

```{r, echo=FALSE}
joint_data <- process_data_mqcv_countries(
  covid_data = data_CV19_monthly,
  mq_data = data_mq,
  inc_countries = "USA",
  exc_countries = NULL,
  product_subset = "personal_protective_equipment", #When NULL, all the products are combined
  covid_subset = "new_cases"
  )
```

### B5.2. Plotting

```{r, echo=FALSE}

inc_countries = "United States of America"
product_subset = "Personal Protective Equipment"
covid_subset = "New COVID-19 cases"


# ##plot-seperately
# plot(joint_data$year_month, joint_data$cv19_variable, type = "l", col = "blue",xlab = "Date", ylab = "COVID-19 cases", main = "COVID-19 cases Over Time")
#
# plot(joint_data$year_month, joint_data$mq_events, type = "l", col = "red",xlab = "Date", ylab = "MQ Events", main = "MQ Events Over Time")



##Plot together
#Calculate the scale factor dynamically
scale_factor <- max(joint_data$cv19_variable, na.rm = TRUE) / max(joint_data$mq_events, na.rm = TRUE) *0.8 #introducing a "visual reduction factor" so MQ evnts line is smaller relative to CV19

  ### Plot correlations over time
  plot = ggplot(joint_data, aes(x = year_month)) +
    geom_line(aes(y = cv19_variable, color = "cv19_cases", group = 1)) +
    geom_line(aes(y = mq_events * scale_factor, color = "MQ Events", group = 1)) +
    scale_y_continuous(
      name = "cv19_cases",
      sec.axis = sec_axis(~./scale_factor, name = "MQ Events")
    ) +
    labs(title = str_c("CV19 ", str_to_title(gsub("_", " ", covid_subset)), " and MQ Events over time \n", " For ", inc_countries), x = "Date", subtitle = str_c(product_subset)) +
    scale_color_manual("",
                       breaks = c("cv19_cases", "MQ Events"),
                       values = c("blue", "red")) +
    theme(
      axis.text.x = element_text (angle = 45, hjust = 1, size = 8),
      axis.title.y.right = element_text(color = "red"),
      axis.title.y.left = element_text(color = "blue")
    ) +
    # scale_x_discrete(limits = unique(category_data$year_month)) +
    geom_point(aes(y = mq_events * scale_factor, color = "MQ Events"),
               size = 2) +
    geom_point(aes(y = cv19_variable, color = "cv19_cases"),
               size = 2)

  print(plot)
```

### B5.3. Cross-correlations (incl stationarity & autocorrelation)

#### Step 0: Define the time series

```{r}
####define the series that needs to be compared
x_series_CV19 <- joint_data$cv19_variable
y_series <- joint_data$mq_events

```

#### Step 1: Stationarity check and differencing

```{r}
# ##Check stationarity x-series
# check_stationarity(x_series_CV19) #non-stationary
# #1st time differencing
# check_stationarity(diff(x_series_CV19)) #stationary

##Check stationarity y-series
check_stationarity (y_series) #stationary

```

```{r, message=FALSE, echo=TRUE}
##################################
### Information for dataframe  ###
##################################

# ##Check stationarity x-series
# stat_check_x <- adf_until_stationary(x_series_CV19)
# print_check_stationarity(stat_check_x)
# #Define stationary x_series
# x_series_CV19_s <- stat_check_x$transformed_series[[1]]


##Check stationarity y-series
stat_check_y <- adf_until_stationary(y_series)
print_check_stationarity(stat_check_y)
#Define stationary y_series
y_series_s <- stat_check_y$transformed_series[[1]]
```

#### Step 2: Autocorrelation analysis using ACF and PACF

```{r}
# ##X-series
# acf_x <- acf(x_series_CV19_s, main= "New COVID-19 cases")
# pacf_x <- pacf(x_series_CV19_s, main= "New COVID-19 cases")

##Y-series
acf_y <- acf(y_series_s, main= product_subset)
pacf_y <- pacf(y_series_s, main= product_subset)
```

No ACF or PACF thus no need to perform prewhitening.

#### Step 3: baseline cross-correlation on stationary data

```{r}
###baseline cross-correlation on stationary data
baseline_ccf <- ccf(x_series_CV19_s, y_series_s, na.action = na.omit,
main = str_c("New COVID-19 cases (x) and ", product_subset, " (y)"), ylab = "baseline CCF")
```

```{r, message=FALSE, echo=FALSE}
##################################
### Information for dataframe  ###
##################################

##Extract the lags that are significant
#significance lines (CCF uses the 95% CI)
sig_line_rccf <- 1.96 / sqrt(baseline_ccf$n.used)

#filter lags above significance level
sig_lags_rccf <- baseline_ccf$acf %>%
  as.numeric() %>%
  tibble(acf = .) %>%
  mutate(lag = baseline_ccf$lag %>% as.numeric()) %>%
  filter(abs(acf) > sig_line_rccf)
sig_lags_rccf


#Range of lags analysed
min_lag <- min(baseline_ccf$lag)
max_lag <- max(baseline_ccf$lag)

#Print results
cat(
  "Significance threshold (95% CI):", sig_line_rccf, "\n",
  "The range of analysis: ", min_lag, " to ", max_lag, "\n",
    "Significant lags:",
  sep = "")
```

#### Step 4: Prewhitened cross-correlation

No ACF/PACF (ánd no cross-correlation detected for the baseline CCF) Thus no need to proceed and perform prewhitening.
